---
title: "R Bridge Week 3 Final Project"
author: "Ilya Kats"
date: "January 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Bridge Course Final Project

Select a data set and analyze. The presentation approach is up to you, but it should contain the following:

1. __Data Exploration:__ This should include summary statistics, means, medians, quartiles, or any other relevant information about the data set. Please include some conclusions in the R Markdown text.
2. __Data Wrangling:__ Please perform some basic transformations. They will need to make sense but could include column renaming, creating a subset of the data, replacing values, or creating new columns with derived data (for example, if it makes sense you could sum two columns together).
3. __Graphics:__ Please make sure to display at least one scatter plot, box plot and histogram. Don't be limited to this. Please explore the many other options in R packages such as ggplot2.
4. __Meaningful Question for Analysis:__ Please state at the beginning a meaningful question for analysis. Use the first three steps and anything else that would be helpful to answer the question you are posing from the data set you chose. Please write a brief conclusion paragraph in R markdown at the end.

## Setup

Load locally saved CSV data file to be used in following tasks. The data file contains environmental conditions in New York City. Data is courtesy of [vincentarelbundock](http://vincentarelbundock.github.io/Rdatasets/).

```{r}
nyc.environment <- read.table(file = "environmental.csv", header = TRUE, sep = ",")
```