---
title: "R Bridge Week 2 Assignment"
author: "Ilya Kats"
date: "January 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load locally saved CSV data file to be used in following tasks. The data file contains environmental conditions in New York City. Data is courtesy of [vincentarelbundock](http://vincentarelbundock.github.io/Rdatasets/).

```{r}
nyc.environment <- read.table(file = "environmental.csv", header = TRUE, sep = ",")
```

## Task 1

Use the summary function to gain an overview of the data set. Then display the mean and median for at least two attributes.

#### Summary
```{r}
summary(nyc.environment)
```

#### Mean and median of temperature
```{r}
mean(nyc.environment$temperature)

median(nyc.environment$temperature)
```

#### Mean and median of radiation
```{r}
mean(nyc.environment$radiation)

median(nyc.environment$radiation)
```

#### Save values for future use
```{r}
# Save temperature values for future use
df.stats <- data.frame(Mean = mean(nyc.environment$temperature), Median = median(nyc.environment$temperature), row.names = "FullSet-Temp")

# Add radiation values for future use 
df.stats <- rbind(df.stats, data.frame(Mean = mean(nyc.environment$radiation), Median = median(nyc.environment$radiation), row.names = "FullSet-Rad"))
```

## Task 2

Create a new data frame with a subset of the columns and rows. Make sure to rename it.

```{r}
# Select temperature and radiation values for 10 random rows, 
# rename columns, sort the selected subset, rename rows 
df.nyc <- data.frame(nyc.environment[sample(1:nrow(nyc.environment), 10), c(1, 3:4)])
names(df.nyc) <- c("orig.row","rad","temp")
df.nyc <- df.nyc[order(df.nyc$orig.row), ]
row.names(df.nyc) <- 1:10
df.nyc
```

## Task 3

Create new column names for the new data frame.

_Note: I have renamed the columns in task 2 before I got to this task, but figured I'll try to find another way. Plus, I like more descriptive full names anyway._

```{r}
df.nyc <- setNames(df.nyc, c("OriginalRow","Radiation","Temperature"))
df.nyc
```

## Task 4

Use the summary function to create an overview of your new data frame. Then print the mean and median for the same two attributes. Please compare.

#### Summary
```{r}
summary(df.nyc)
```

#### Mean and median of temperature
```{r}
mean(df.nyc$Temperature)

median(df.nyc$Temperature)
```

#### Mean and median of radiation
```{r}
mean(df.nyc$Radiation)

median(df.nyc$Radiation)
```

#### Save new values for comparison
```{r}
# Add subset temperature values
df.stats <- rbind(df.stats, data.frame(Mean = mean(df.nyc$Temperature), Median = median(df.nyc$Temperature), row.names = "SubSet-Temp"))

# Add subset radiation values 
df.stats <- rbind(df.stats, data.frame(Mean = mean(df.nyc$Radiation), Median = median(df.nyc$Radiation), row.names = "SubSet-Rad"))

df.stats
```

__Comparison:__ Mean and median values between a full data set and a subset are obviously different, but since a subset represents a random sample, they should be relatively close. Of course, since the sample is small there may be exceptions to this. Also, I was slightly unsure whether you want descriptive comparison or an actual _if...then_ comparison with a printout. As you see, I went with descriptive a comparison/analysis.

## Task 5

For at least 3 values in a column please rename so that every value in that column is renamed. For example, suppose I have 20 values of the letter "e" in one column. Rename those values so that all 20 would show as "excellent."

_Please note that my data only had numeric values, which were not well suited for this task. I have decided to add a character column based on my data to do these data manipulations. The solution may include more than was requested, but it was great practice for me._

```{r}
# Add a column to hold Beaufort wind force scale with possible values applicable to the data set
nyc.environment <- cbind(nyc.environment, beaufort = factor(NA, levels = c("la", "lb", "gb", "mb", "fb")))

# Populate Beaufort scale based on wind speed
nyc.environment[nyc.environment$wind <= 3, "beaufort"] <- "la"
nyc.environment[nyc.environment$wind > 3 & nyc.environment$wind <= 7, "beaufort"] <- "lb"
nyc.environment[nyc.environment$wind > 7 & nyc.environment$wind <= 12, "beaufort"] <- "gb"
nyc.environment[nyc.environment$wind > 12 & nyc.environment$wind <= 20, "beaufort"] <- "mb"
nyc.environment[nyc.environment$wind > 20, "beaufort"] <- "fb"

# Quick data check
head(nyc.environment)

# Convert Beaufort scale column from factor to character in order to add new values previously not included
v.beaufort <- as.character(nyc.environment[,"beaufort"])

# Replace old values with more descriptive new values
v.beaufort[v.beaufort == "la"] <- "light air"
v.beaufort[v.beaufort == "lb"] <- "light breeze"
v.beaufort[v.beaufort == "gb"] <- "gentle breeze"
v.beaufort[v.beaufort == "mb"] <- "moderate breeze"
v.beaufort[v.beaufort == "fb"] <- "fresh breeze"

# Convert back to a factor and replace old factor with new one in the data set
nyc.environment[,"beaufort"] <- as.factor(v.beaufort)
```

## Task 6

Display enough rows to see examples of all of steps 1-5 above.

_Please note that for tasks 1 through 4 results were displayed within each task, so the sample of the data below is mostly to demonstrate solution to task 5._ 

```{r}
nyc.environment[1:15, ]
```

## Task 7 (Bonus)

Place the original .csv in a github file and have R read from the link.

```{r}
require(RCurl)

# Load data file from GitHub
git.nyc.environment <- read.csv(text=getURL("https://raw.githubusercontent.com/ilyakats/CUNY-R-Bridge-Workshop/master/git-environmental.csv"), header = TRUE, sep = ",")

# Quick data check
head(git.nyc.environment)
```

## Cleanup

```{r}
rm(nyc.environment, git.nyc.environment, df.nyc, df.stats)
```